---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## load libraries

```{r}
library(ggplot2)

```



## Loading and preprocessing the data

### Data reading

```{r}
Data=read.csv('activity.csv')
summary(Data)
head(Data)
```


## What is mean total number of steps taken per day?

```{r}
TotalSteps<-tapply(Data$steps, Data$date, sum, na.rm=TRUE)
TotalSteps
hist(TotalSteps, main = "Total number of steps taken per day", xlab = "Total steps taken per day", col = "Red")
```

### Mean
```{r}
mean(TotalSteps)
```
### Median
```{r}
median(TotalSteps)
```


## What is the average daily NewData pattern?

```{r}
MeanStepsInt<-tapply(Data$steps, Data$interval, mean, na.rm=TRUE)
plot(names(MeanStepsInt), MeanStepsInt, type = "l", col="darkblue", lwd = 2, xlab="Interval",
     ylab="Average number of steps", main="Average number of steps per intervals")

```

###  the maximum number of steps
```{r}
M<-max(MeanStepsInt)
as.numeric(names(MeanStepsInt[MeanStepsInt==M]))
```

## Imputing missing values
### the total number of missing values
```{r}
sum(is.na(Data$steps))
```
### New dataset
```{r}
imputed_steps <- MeanStepsInt[match(Data$interval,names(MeanStepsInt))]
NewData<-transform(Data, steps = ifelse(is.na(Data$steps), yes = imputed_steps, no = Data$steps))

```

### Histogram

```{r}
TotalSteps2<-tapply(NewData$steps, NewData$date, sum)
TotalSteps2
hist(TotalSteps2, main = "Total number of steps taken per day", xlab = "Total steps taken per day", col ='green')
```

### Mean
```{r}
mean(TotalSteps2)
```
### Median
```{r}
median(TotalSteps2)
```

The mean didn’t change after the replacements of NAs, the median changed about 0.1% of the original value.

## Are there differences in NewData patterns between weekdays and weekends?

```{r}
NewData$date <- as.Date(strptime(NewData$date, format="%Y-%m-%d"))
NewData$datetype <- sapply(NewData$date, function(x) {
        if (weekdays(x) == "sábado" | weekdays(x) =="domingo") 
                {y <- "Weekend"} else 
                {y <- "Weekday"}
                y
        })
```

```{r}
NewData_by_date <- aggregate(steps~interval + datetype, NewData, mean, na.rm = TRUE)
plot<- ggplot(NewData_by_date, aes(x = interval , y = steps, color = datetype)) +
       geom_line() +
       labs(title = "Average daily steps by type of date", x = "Interval", y = "Average number of steps") +
       facet_grid(datetype~.)
print(plot)

```




